### 2013-12-15 paper/paper.tex
LaTeX version. Also did some more revision.

\documentclass{article}

\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{fancyhdr}

\usepackage{hyperref}
\setlength{\headheight}{15pt}
\pagestyle{fancyplain}

\setlength{\parindent}{0pt}
\setlength{\parskip}{2ex}

\tikzstyle{int}=[draw, minimum size=4em]
\tikzstyle{init} = [pin edge={to-,thin,black}]
\lhead{Dynamic Hover Detection: Creating GUIs in Elm}
\rhead{Goldstein \& Alter}

\makeatletter
\renewcommand{\maketitle}{ % Customize the title - do not edit title and author name here, see the TITLE block below
    \begin{flushright} % Right align
        {\LARGE\@title} % Increase the font size of the title

        \vspace{40pt} % Some vertical space between the title and author name

        {\large\@author} % Author name

        \vspace{30pt} % Some vertical space between the author block and abstract
    \end{flushright}
}

% This begins the document
\title{\textbf{Dynamic Hover Detection:\\Creating GUIs in Elm}}
\author{Max Goldstein \& Eliot Alter}

\begin{document}
\thispagestyle{empty} % No header on first page
\maketitle

\begin{abstract}
The language Elm was designed to be a practical application of
Functional Reactive Programming (FRP) to create Graphical User
Interfaces (GUIs). We attempted to create a menu bar as commonly seen on
desktop operating systems, and quickly discovered an apparent
incompatibility between Elm's mouse-hover detection primitive and the
constraints of the language. This concern only surfaces when dealing
with highly dynamic data. We present a non-obvious technique to address
the issue without modifying the Elm compiler. It generalizes to other
functions in Elm's \texttt{Graphics.Input} library, which includes
(besides hover detection) GUI mainstays such as buttons, checkboxes, and
text fields. We also contribute other techniques for the representation and
display of menus in Elm, and contrast our work with an existing Elm web
application.
\end{abstract}

\section{Introduction}\label{introduction}

Elm was introduced in March 2012 in Evan Czaplicki's senior thesis,
\emph{Concurrent FRP for GUIs}. He and his adviser, Stephen Chong,
published \emph{Asynchronous FRP for GUIs} at PLDI 2013. Both papers are
comprehensive overviews of Elm, and additionally provide excellent
literature reviews of previous FRP GUI endeavors, of which there are
several. Elm compiles down to JavaScript to run in the browser, making
it remarkably portable.

We chose menus as an example GUI to implement in Elm. The particular
design places the top-level menu at the top of the screen, similar to
Mac OS X and many Linux distributions, with selections coming down from
the top. Menus extend when the top-level item is hovered upon, and
remain extended while the mouse hovers over any item in the menu.
Therefore it is necessary to know hover information about each menu
item. This time-varying information is also used to detect selections
upon click and highlight the moused-over item. It is simple to do this
when the hover-detecting area is constant. Our paper describes the much
more difficult task of managing time-varying hover information about
time-varying areas.

There are two features of Elm we are deliberately avoiding. First is the
extensive raster drawing library, \texttt{Graphics.Collage}. Dynamic
hover detection is not problematic when using this library because it
can be done purely through manual collision detection. However, if we
used this library, our GUI would be a single raster animation and not a
DOM tree. Secondly, the \texttt{Graphics.Input} library contains
wrappers around HTML checkboxes and dropdowns. We consider creating GUIs
with these to be uninteresting and wanted to pursue something more
general. This led to the choice of an interface typically found in the
operating system rather than the browser. We do refer to the native GUI
constructs to show how the technique we develop for hovering
generalizes, as the API for buttons and hover detection is very similar.

It is difficult for the authors to assess what level of knowledge should
be assumed on the part of the reader. Firstly, readers will range from
Elm's creators and experienced users, who are already familiar with its
restrictions and standard libraries, to those with no FRP background,
who need an introduction to signals. Secondly, while we have found no
prior discussion on either the problem we identify nor its solution, we
cannot know for certain that either are novel. We continue in the hopes
of presenting new and non-trivial techniques to the Elm community.

We contribute:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  The identification of an apparent limitation in Elm's hover detection
  library, solved by a non-trivial usage pattern that does not require
  modifications to the Elm compiler or runtime, and that generalizes to
  other functions in \texttt{Graphics.Input}.
\item
  An implementation of desktop-style menu in Elm, which incorporates
  several noteworthy ``tricks'' as well as general style practices.
\item
  An analysis of TodoFRP, the current state-of-the-art in dynamic Elm
  GUIs. We demonstrate how it operates in the absence of our technique,
  and how it could operate in its presence.
\end{itemize}

Section 2 introduces Elm, signals, and the prohibition on signals of signals. It
is targeted to readers familiar with functional programming but not FRP, and may
be skipped by those already familiar with Elm. Section 3 presents Elm's hover
detection for DOM elements, why it initially appears inadequate, and how it can
be extended to meet our needs. Section 4 explains our menu implementation in
detail. Section 5 analyzes TodoFRP. Section 6 concludes with a notice to the Elm
community.

\section{Elm for Functional
Programmers}\label{elm-for-functional-programmers}

A typical functional program is \emph{transformative}: all input is
available at the start of execution, and after a hopefully finite amount
of time the program terminates with some output. In contrast, Elm
programs are \emph{reactive}: not all input is available immediately and
the program may indefinitely adjust output with each input. In simple
programs, the inputs at a given time fully determine the output. More
complex programs will take advantage of Elm's ability to remember state.

\subsection{Signals: Time-varying
values}\label{signals-time-varying-values}

A time-varying value of a polymorphic type \texttt{a} is represented by
\texttt{Signal a}. For example, the term \texttt{constant 150} has type
\texttt{Signal Int}. The combinator \texttt{constant} creates a signal
whose value never changes. A more interesting signal is the primitive
\texttt{Window.dimensions : Signal (Int, Int)}. This signal represents
the browser window size and updates whenever it is resized. Signals are
asynchronous in that they update at no set time, just as the window may
remain the same size idefinitely. Signals update in discrete events, but
are continuous in the sense that they are always defined.

The function \texttt{lift} allows us to execute a pure function on a
signal of inputs, producing a signal of outputs. (Although lifting is a
general functional concept, in Elm it has only this meaning.)

\begin{verbatim}
lift : (a -> b) -> Signal a -> Signal b
\end{verbatim}

For example, we can multiply the width and height of the window together
to find its area.

\begin{verbatim}
area : Signal Int
area = lift (uncurry (*)) Window.dimensions
\end{verbatim}

In this case, the lifted function is multiplication, uncurried as to
operate on pairs. As the Window library exposes width and height as both
a pair and individually, we can also write
\texttt{area2 = lift2 (*) Window.width Window.height}.

We can print the current area to the screen with \texttt{main = lift asText
area}. The primitive \texttt{asText : a -\textgreater{} Element} renders almost
anything into an Element, which represents a DOM element.\footnote{Those
    following along in an Elm compiler, such as the one available at
    \texttt{elm-lang.org/try}, should add \texttt{import Window} to the top of
the file.}

The signals in an Elm program can be thought of as a Directed Acyclic
Graph (DAG). Many signals depend on other signals for their output. For
example, the \texttt{area} signal depends on \texttt{Window.dimensions}
signal. Similarly, the \texttt{area2} signal depends on both the values
of the \texttt{Window.width} and the \texttt{Window.height} signals.
When an event is fired on a signal, it propogates down the DAG to all
signals who depend on that signal and the outputs of those signals are
reevaluated.

\subsection{Remembering State}\label{remembering-state}

Signals can remember state by using the \texttt{foldp} combinator.
Familiar list folds apply a binary operation of an element and an
accumulator to produce a new accumulator, over each list element in
sequence. Folding from the \emph{past} operates on all of the values of
a signal as they occur and produces a signal of the accumulator.

\begin{verbatim}
foldr : (a -> b -> b) -> b -> [a]      -> b
foldp : (a -> b -> b) -> b -> Signal a -> Signal b
\end{verbatim}

When the event signal updates, a pure function is called with the new
event and the old accumulator (a default is supplied), producing a new
accumulator that is the new value of the output signal. For example,
\texttt{foldp max 0 Window.width} is a signal of the maximum width ever
obtained by the window. With \texttt{foldp}, it is possible to create
signals that depend on every event to ever occur on a signal. However,
most folded functions do not store every value explicitly (cons is an
exception) and space can be saved by remembering only the accumulator.

\subsection{No Signals of Signals}\label{no-signals-of-signals}

The ability to create a signal dependent on past state has a potentially
disastrous implication for space performance. Czaplicki and Chong
explain the preventive measure and the problem:

\begin{quote}
The type system rules out programs that use signals of signals, for the
following reason. Intuitively, if we have signals of signals, then after
a program has executed for, say 10 minutes, we might create a signal
that (through the use of \texttt{foldp}) depends on the history of an
input signal, say \texttt{Window.width}.
\end{quote}

That is, we can define a clock that measures the amount of time since
program execution began:

\begin{verbatim}
clock : Signal Time
clock = foldp (+) 0 (fps 10)
\end{verbatim}

The \texttt{fps n} combinator produces a signal of the time elapsed
since its last event, updated \texttt{n} times a second. We sum these
deltas starting with zero. Then we create a function from a time to one
of two signals, one of which is trivial and one of which depends on the
history of \texttt{Window.width}:

\begin{verbatim}
switcher : Time -> Signal Int
switcher t = if t < 10*minute
             then constant 0
             else foldp max 0 Window.width
\end{verbatim}

We could, were the type system not principled on disallowing it, create

\begin{verbatim}
switched : Signal (Signal Int)
switched = lift switched clock
\end{verbatim}

Why is this problematic? Czaplicki and Chong continue,

\begin{quote}
To compute the current value of this signal, should we use the entire
history of \texttt{Window.width}? But that would require saving all
history of \texttt{Window.width} from the beginning of execution, even
though we do not know whether the history will be needed later.
Alternatively, we could compute the current value of the signal just
using the current and new values of \texttt{Window.width} (i.e.,
ignoring the history). But this would allow the possibility of having
two identically defined signals that have different values, based on
when they were created. We avoid these issues by ruling out signals of
signals.
\end{quote}

That is, in order to have both state and signals of signals, the program
must either remember indefinitely every event to ever occur, so that a
newly created signal can use them, or tolerate signals that vary based
only on when they were created, losing referential transparency. Elm's
creators declined either option and disallowed signals of signals.

To a reader familiar with Haskell, this means signals are functors (and
in fact applicative functors) but not monads, as monads support the
following operation:

\begin{verbatim}
join :: Monad m => m (m a) -> m a
\end{verbatim}

Such an operation for signals would condense a
\texttt{Signal (Signal a)} into a mere \texttt{Signal a}, but it cannot
exist in general.

\section{Detecting Hover Information}\label{detecting-hover-information}
A cursory inspection of \texttt{Graphics.Input} leads to the conclusion that
menus cannot be implemented as we had hoped. The remarkable solution generalizes
from detecting hover information to other GUI functions in the library.

\subsection{A Na\"{i}ve Menu}\label{a-naive-menu}

Elm provides a function to obtain the hover information from an Element.

\texttt{hoverable : Element -\textgreater{} (Element, Signal Bool)}

The returned Element is visually identical to argument, but now detects
hover information. The signals of Booleans reflects the hover status of
the this Element, not the original. This function works well when the
Element is pure (not a signal). For example, if the top-level menu items are
known statically and never change, they can be rendered as pure Elements. The
\texttt{hoverable} function allows us to create Elements for their submenus.
However, if we try to lift the function, we get

\texttt{lift hoverable : Signal Element -\textgreater{} Signal (Element, Signal Bool)}

Although it may not appear to be so at first, the result type is a signal of
signals. It is possible to transform the result into \texttt{Signal Signal
(Element, Bool)}. A more mentally convenient type, though impossible to obtain,
is \texttt{(Signal Element, Signal Signal Bool)}. There does not exist a general
join function to operate on the \texttt{Signal Signal Bool}.

This is rather unfortunate, as we need to detect hover information on a dynamic
element for a number of reasons. First, after a menu is extended by hovering
over its parent, it must remain so as long as the mouse hovers over the menu
itself. Secondly, if we wish the indicate the currently hovered-over item to the
user by visually highlighting it, the element must be dynamic. Thirdly, we do
not wish to require that menu labels be string literals in the source code,
which would be necessary if we could not use signals of elements. Rather they
may be retrieved or generated in some other manner, or they may change as the
user switches among applications. Finally, in order to determine when a user has
clicked on an Element so that the system may respond, we must know which one is
being hovered over.\footnote{Once this is known, \texttt{sampleOn Mouse.clicks}
can be applied to obtain a signal of the hovered-over element updated on each
click.}

\subsection{Dynamic Hovering}
However, it is possible to implement the following function:

\texttt{hoverableJoin: Signal Element -\textgreater{} (Signal Element, Signal Bool)}

It is done using the more general \texttt{hoverables} (note the plural)
primitive, of the following type:

\begin{verbatim}
hoverables : a -> { events : Signal a,
                    hoverable : (Bool -> a) -> Element -> Element }
\end{verbatim}

The polymorphic \texttt{a} type can serve as an identifier. The first
value supplies the default value of \texttt{events} (signals must always
be defined and so a default value is required). The returned record
includes the \texttt{events} signals and the \texttt{hoverable}
function, which in general may be applied multiple times so that
multiple elements report on \texttt{events}. The
\texttt{(Bool -\textgreater{} a)} is used to identify which Element
experienced the event; a common use is \texttt{a = (Int, Bool)} where
the integer identifies the Element and the Bool is the event.
Additionally, \texttt{hoverables} is used to implement
\texttt{hoverable}:

\begin{verbatim}
hoverable : Element -> (Element, Signal Bool)
hoverable elem =
    let pool = hoverables False
        in  (pool.hoverable id elem, pool.events)
\end{verbatim}

It ignores the polymorphism (\texttt{a = Bool}) and instead create a
Boolean signal that is originally false and use the identity function to
not alter the hoverable information. With a simple change, we can create
a function that acts on \texttt{Signal Element} instead of
\texttt{Element}:

\begin{verbatim}
hoverablesJoin : Signal Element -> (Signal Element, Signal Bool)
hoverablesJoin elem =
    let pool = hoverables False
    in (lift (pool.hoverable id) elem, pool.events)
\end{verbatim}

Notice that \texttt{pool.hoverable} is partially applied to \texttt{id}
purely, and then lifted on to the argument. This is possible,
utlimately, because \texttt{pool.hoverable} is pure.

Just how novel is this small, but hugely significant change? The
\href{http://docs.elm-lang.org/library/Graphics/Input.elm\#hoverables}{documentation}
for \texttt{hoverables} states that it allows users to ``create and
destroy elements dynamically and still detect hover information,'' but
gives no further indicators on how to do so. Though the Elm website is
full of examples, there are none for either \texttt{hoverable} or
\texttt{hoverables}. Moreover, in the
\href{https://groups.google.com/d/msg/elm-discuss/QgowLy5jdhA/CZQfjkbjMsEJ}{mailing
list post} that introduced these functions, Czaplicki said that
``\texttt{hoverables} is very low level, but the idea is that you can
build any kind of nicer abstraction on top of it.'' We have done just
that.

When a element changes, \texttt{hoverablesJoin} attaches the same
hovering signal to the new element. If we were to display both the old
and the new elements on the screen, hovering over either would trigger
the signal.

More dangerously, it becomes easy to create an infinite loop. Suppose an
Element shrinks on hover. Suppose the cursor hovers on the Element,
which is then replaced by a smaller Element, so the cursor is no longer
hovering on the Element. Then the original Element is put back, but is
now being hovered on! This condition manifests itself as flickering
between the two Elements. It is unavoidable in any language or system
that allows hover targets to change size in response to hover events.

\subsection{Generalized Joins on \texttt{Graphics.Input}}

It is not just hover detection that follows the convention of the two
names primitives, singular and plural, but most of Elm's wrappers around
HTML GUI components. They are therefore receptive of the same join
technique. For labelled buttons, press events are represented by unit or
a provided identifier:

\begin{verbatim}
 button : String -> (Element, Signal ())
 buttons : a -> { events : Signal a,
                  button : a -> String -> Element }
\end{verbatim}

If we wish to create a button whose string label varies, the types allow
the same exact same technique to be applied. (We have not tested whether
this approach works, but the types allow it.) The same goes for text
fields, which can be given text based on what else is known in the
program, and checkboxes. As currently implemented, checkboxes have
Boolean state so dynamic behavior has limited use. However, we can
imagine the ability to disable (``gray-out'') a checkbox in response to
other choices in a form.

Drop-down menus do not use the same paradigm as the rest; there is
\texttt{dropDown} but no \texttt{dropDowns}. This is unfortunate not
only because there seems to be no way to supply a
\texttt{Signal {[}String{]}}, but because drop-downs are commonly used
in groups. Multiple menus reporting on one signal would be very
convenient. If the menus were dynamic, it may be useful to determine
when an event occurs because the user makes a selection vs.~when the
menu updates and the previously selected string is no longer available.

% These are likely unnecessary and cause page breaks in bad places.
%
%\texttt{Graphics.Input} does not currently support radio buttons.
%
%As an aside, \texttt{button} is an impure function. One can create two
%buttons with the same label and place them in different places on the
%screen (or place one not at all), and the signals of press events will
%be different. Lifting a joined version onto a signal of strings does not
%change this fact.

\section{Implementing Menus}\label{implementing-menus}

A menu can be thought of as an instance of a tree.

\texttt{data Tree a = Tree a [a]}

Using \texttt{hoverablesJoin}, we created a structure of type
\texttt{Tree (Signal Element, Signal Bool)} to represent our menus.
Using the combinator
\texttt{combine : {[}Signal a{]} -\textgreater{} Signal {[}a{]}}, we
were able to turn this structure into a
\texttt{Signal (Tree (Element, Bool))}. Once we had that structure, we
could render the menu using a pure function from
\texttt{Tree (Element, Bool) -\textgreater{} Element} and then lift it.
Similarly to limiting the use of the IO monad in Haskell code, it is
advantageous to limit the use of signals in Elm code.

Our first trick was to replace submenus not shown on the screen with
spacers. A spacer is simply a blank rectangle with a specified with and
height
(\texttt{spacer : Int -\textgreater{} Int -\textgreater{} Element}). For
every menu item, we create both the menu Element and a spacer. The
spacer has the same width as the Element, but a height of 1. We then
switch between menu and the spacer based on hover information. This
helped in two ways. First of all, it allowed us to choose between two
options without creating a signal of signals. Secondly, the spacer
helped to align the submenu with its parent element.

Our second trick fixed a race condition. A submenu was rendered to the
screen if one of three conditions was met: the mouse was hovering upon
its parent, the mouse was hovering upon it, or the mouse was hovering
upon any of its descendents. As the mouse moved from the parent to the
child, the child was often replaced with a spacer before it could detect
that the mouse was hovering over it. To fix this problem, we created the
following function:

\begin{verbatim}
extend : Signal Bool -> Signal Bool
extend b = lift2 (||) b (delay millisecond b)
\end{verbatim}

Using \texttt{extend} makes a \texttt{Signal Bool} wait a millisecond to
transition from True to False. By applying this function to every
hovering Boolean in the Menu structure, we gave the mouse time to move
from a menu Element to its submenu before that submenu disappeared.

Interestingly enough, we were able to implement menus without explicitly
storing the state of the menu. There are two ways to remember state in
Elm code: \texttt{foldp}, or Elm's implementation of Arrowized FRP known
as \texttt{Automaton}. However, in our implementation we use the current
browser DOM state to remember our state. Instead of storing the menu
currently being displayed on the screen, we can derive it based on which
Element (if any) the mouse is currently hovering over. The elimination
of state from our code allows it to be much cleaner.

Note that the input to our menu is a \texttt{Tree (Signal String)}, but
not a \texttt{Signal (Tree String)}. We are able to map from
\texttt{Signal Element -\textgreater{} (Signal Element, Signal Bool)},
but not from \texttt{Element -\textgreater{} (Element, Bool)}.
Therefore, if we tried to turn a \texttt{Signal (Tree String)} into a
\texttt{Signal (Tree (Element, Bool))}, we would need to lift an
impossible hoverable function of type
\texttt{Element -\textgreater{} (Element, Bool)} on each element.
Therefore, each individual menu element can contain dynamic information,
but the menu structure must be static. We can, in practice, get around
this restriction by creating a tree that is larger than necessary and
filling the unused nodes with empty string, which our implementation
handles appropriately.

A possible solution for that issue would be to write a function of type
\texttt{Signal (Tree a) -\textgreater{} Tree (Signal a)}. Although the
\texttt{combine} combinator can turn a list of signals into a signal of
lists, there isn't an inverse to turn a signal of lists into a list of
signals. It is impossible to write the function
\texttt{split : Signal {[}a{]} -\textgreater{} {[}Signal a{]}} because
the size of the list in the input is dynamic, whereas the size of the
output list is static. Since the \texttt{Tree} data structure contains
lists, it is therefore impossible to write a function of type
\texttt{Signal (Tree a) -\textgreater{} Tree (Signal a)}.

\section{Related Work: TodoFRP}\label{related-work-todofrp}

We examine \href{https://github.com/evancz/TodoFRP}{TodoFRP}, a simple
Elm web app created by Czapliki, to both pinpoint the problem and show
its universality. TodoFRP is the current state-of-the-art in highly
reactive Elm GUIs. It provides examples of different levels of
reactivity, a familiar context for Elm veterans, and a few dirty tricks
of its own.

TodoFRP presents the user with a text field asking, ``what needs to be
done?''. Entered todo entries become DOM elements, which can be deleted
with a ``x'' button, also a DOM element. The button is implemented using
the \texttt{Graphics.Input} function

\begin{verbatim}
customButtons : a -> { events : Signal a,
                       customButton : a -> Element -> Element -> Element -> Element }
\end{verbatim}

Notice the similarity with \texttt{hoverables}. Each call of
\texttt{customButton} provides the identifier event when the button is
clicked, and three (pure) Elements to display: one normally, one on
hover, and one on click. The result is an \texttt{Element}, not a
\texttt{Signal Element}, that nevertheless changes among those three in
response to the mouse. This is possible because the result Element's
dimensions are taken to be the maximum of the three inputs' dimensions.
Even if the Elements have different sizes, the resulting element and
therefore the hover surface remains fixed in size. Although the same
join technique can be applied, we find it likely that it would not work
as intended.

In the case of TodoFRP, these Elements are diferent colors of the ``x''
and the same for each todo entry. The polymorphic \texttt{a}s are unique
identifiers (ascending integers) for each entry.

The todo label Elements are dynamic and do not detect hover information.
The button Elements that do detect hover information are known
statically. Without \texttt{hoverablesJoin}, it would be impossible to
dismiss a todo by clicking on its text label.

\section{Conclusion: To the Elm
Community}\label{conclusion-to-the-elm-community}

It's true that we've used the \texttt{hoverables} function in a way that
it was (probably) never meant to be used, and there are some caveats
involved in doing so. Many small GUIs do not require dynamic,
hover-detecting elements. However, most large mouse-based GUIs do, and
creating them in Elm will necessarily encounter the obstacles we have
described.

We have implemented all of this without language modifications. However,
it is hoped that as the community becomes more familiar with functional
GUIs, new libraries are added to incorporate some of our tricks, or even
make them unnecessary. Elm's upcoming third-party library sharing system
looks to be an excellent opportunity to refine abstractions and idioms
for GUIs.

We present these techniques and analysis in the hopes they are of
service to the Elm and FRP communities. We hope this paper becomes
useful in Elm's goal of making GUIs simpler to implement and more robust
to use.

\subsection{Acknowledgements}\label{acknowledgements}

We would like to thank Evan Czaplicki and Stephen Chong for creating
Elm, and the Elm community for growing it. We thank Norman Ramsey for
his guidance through functional programming, and our paper reviewers,
\ldots{} .

\subsection{References}\label{references}
\end{document}

### 2013-12-15 paper/paper.tex
Mostly spelling fixes

@@
 asynchronous in that they update at no set time, just as the window may
-remain the same size idefinitely. Signals update in discrete events, but
+remain the same size indefinitely. Signals update in discrete events, but
 are continuous in the sense that they are always defined.
@@
 of the \texttt{Window.width} and the \texttt{Window.height} signals.
-When an event is fired on a signal, it propogates down the DAG to all
+When an event is fired on a signal, it propagates down the DAG to all
 signals who depend on that signal and the outputs of those signals are
@@
 purely, and then lifted on to the argument. This is possible,
-utlimately, because \texttt{pool.hoverable} is pure.
+ultimately, because \texttt{pool.hoverable} is pure.
 
@@
 HTML GUI components. They are therefore receptive of the same join
-technique. For labelled buttons, press events are represented by unit or
+technique. For labeled buttons, press events are represented by unit or
 a provided identifier:
@@
 in groups. Multiple menus reporting on one signal would be very
-convenient. If the menus were dynamic, it may be useful to determine
-when an event occurs because the user makes a selection vs.~when the
-menu updates and the previously selected string is no longer available.
+convenient. If the menus were dynamic, it would be useful to determine
+when an event occurred because the user makes a selection vs.~when the
+menu updated and the previously selected string was no longer available.
 
@@
 its parent, the mouse was hovering upon it, or the mouse was hovering
-upon any of its descendents. As the mouse moved from the parent to the
+upon any of its descendants. As the mouse moved from the parent to the
 child, the child was often replaced with a spacer before it could detect
@@
 TodoFRP presents the user with a text field asking, ``what needs to be
-done?''. Entered todo entries become DOM elements, which can be deleted
+done?''. Entered TODO entries become DOM elements, which can be deleted
 with a ``x'' button, also a DOM element. The button is implemented using
@@
 as intended.
+%Note to Max: I think it does work...
 
-In the case of TodoFRP, these Elements are diferent colors of the ``x''
-and the same for each todo entry. The polymorphic \texttt{a}s are unique
+In the case of TodoFRP, these Elements are different colors of the ``x''
+and the same for each TODO entry. The polymorphic \texttt{a}s are unique
 identifiers (ascending integers) for each entry.
 
-The todo label Elements are dynamic and do not detect hover information.
+The TODO label Elements are dynamic and do not detect hover information.
 The button Elements that do detect hover information are known
 statically. Without \texttt{hoverablesJoin}, it would be impossible to
-dismiss a todo by clicking on its text label.
+dismiss a TODO by clicking on its text label.
 
@@
 \subsection{References}\label{references}
+%Make sure to put a link to our GitHub
+
 \end{document}

### 2013-12-15 paper/paper.tex
More revisions

@@
 
-Section 2 introduces Elm, signals, and the prohibition on signals of signals. It
-is targeted to readers familiar with functional programming but not FRP, and may
-be skipped by those already familiar with Elm. Section 3 presents Elm's hover
-detection for DOM elements, why it initially appears inadequate, and how it can
-be extended to meet our needs. Section 4 explains our menu implementation in
-detail. Section 5 analyzes TodoFRP. Section 6 concludes with a notice to the Elm
-community.
+Section 2 introduces Elm, signals, state, and the prohibition on signals of
+signals. It is targeted to readers familiar with functional programming but not
+FRP, and may be skipped by those already comfortable with Elm. Section 3
+presents Elm's hover detection for DOM elements, why it initially appears
+inadequate, and how it can be extended to meet our needs. Section 4 explains our
+menu implementation in detail. Section 5 analyzes TodoFRP. Section 6 concludes
+with a notice to the Elm community.
 
@@
 
-Elm provides a function to obtain the hover information from an Element.
+Elm's \texttt{Graphics.Input} library provides a function to obtain the hover
+information from an Element.
 
@@
 The returned Element is visually identical to argument, but now detects
-hover information. The signals of Booleans reflects the hover status of
+hover information. The signal of Booleans reflects the hover status of
 the this Element, not the original. This function works well when the
 Element is pure (not a signal). For example, if the top-level menu items are
-known statically and never change, they can be rendered as pure Elements. The
-\texttt{hoverable} function allows us to create Elements for their submenus.
-However, if we try to lift the function, we get
+known statically and never change, they can be rendered as pure Elements with
+this function. We can also create a function, perhaps of type \texttt{Bool ->
+Element}, which we can lift onto the hover data to display the submenu only on
+mouseover. The result is a value of type \texttt{Signal Element}. If we try to
+lift \texttt{hoverable} to accept this value, we get
 
@@
 signals. It is possible to transform the result into \texttt{Signal Signal
-(Element, Bool)}. A more mentally convenient type, though impossible to obtain,
+(Element, Bool)}. A more mentally convenient type, though impossible to
+obtain,\footnote{We explain why in Section 4.}
 is \texttt{(Signal Element, Signal Signal Bool)}. There does not exist a general
@@
 over its parent, it must remain so as long as the mouse hovers over the menu
-itself. Secondly, if we wish the indicate the currently hovered-over item to the
-user by visually highlighting it, the element must be dynamic. Thirdly, we do
-not wish to require that menu labels be string literals in the source code,
-which would be necessary if we could not use signals of elements. Rather they
-may be retrieved or generated in some other manner, or they may change as the
-user switches among applications. Finally, in order to determine when a user has
-clicked on an Element so that the system may respond, we must know which one is
-being hovered over.\footnote{Once this is known, \texttt{sampleOn Mouse.clicks}
-can be applied to obtain a signal of the hovered-over element updated on each
-click.}
+itself. Without this information, the menu will disappear as soon as we try to
+mouse over it. Secondly, if we wish the indicate the currently hovered-over item
+to the user by visually highlighting it, the element must be dynamic. Thirdly,
+we do not wish to require that menu labels be string literals in the source
+code, which would be necessary if we could not use signals of elements. Rather,
+they may be retrieved or generated in some other manner, or they may change as
+the user switches among applications. Finally, in order to determine when a user
+has clicked on an Element so that the system may respond, we must know which one
+is being hovered over.\footnote{Once this is known, \texttt{sampleOn
+Mouse.clicks} can be applied to obtain a signal of the hovered-over element
+updated on each click.}
 
@@
 
-It is done using the more general \texttt{hoverables} (note the plural)
-primitive, of the following type:
+Comparing its type to the mentally convenient type of \texttt{lift hoverables},
+we see it is identical except that the Boolean value is a signal, not a signal
+of signals. This function is implemented using the more general
+\texttt{hoverables} (note the plural) primitive, of the following type:
 
@@
 
-It is not just hover detection that follows the convention of the two
-names primitives, singular and plural, but most of Elm's wrappers around
-HTML GUI components. They are therefore receptive of the same join
-technique. For labeled buttons, press events are represented by unit or
-a provided identifier:
+It is not just hover detection that follows the paradigm of two primitives
+with singular and plural names. Most of Elm's wrappers around HTML GUI
+components do so as well, and have analogous types. They are therefore receptive
+of the same join technique. For text-labeled buttons, press events are
+represented by unit or a provided identifier:
 
@@
 in groups. Multiple menus reporting on one signal would be very
-convenient. If the menus were dynamic, it would be useful to determine
-when an event occurred because the user makes a selection vs.~when the
+convenient. If the menus were dynamic, it would be useful for the API to
+indicate when an event occurred because the user makes a selection vs.~when the
 menu updated and the previously selected string was no longer available.
 
-% These are likely unnecessary and cause page breaks in bad places.
-%
-%\texttt{Graphics.Input} does not currently support radio buttons.
-%
-%As an aside, \texttt{button} is an impure function. One can create two
-%buttons with the same label and place them in different places on the
-%screen (or place one not at all), and the signals of press events will
-%be different. Lifting a joined version onto a signal of strings does not
-%change this fact.
+\texttt{Graphics.Input} does not currently support radio buttons, but one could
+imagine them being added in a way similar to the others. However, it is worth
+observing that the pattern itself leaves something to be desired, namely, that
+the functions are impure. One can apply \texttt{hoverable} to an Element twice,
+place the resulting Elements in in different places on the screen (or place one
+not at all), and the signals of Booleans will be different. This is true,
+analogously, of all the functions of similar type in the library. We know of no
+alternative that has been proposed, and do not have one ourselves. However, we
+point out this deficiency in order to illustrate that the library still has room
+for incremental and perhaps even radical growth.
 

### 2013-12-15 paper/paper.tex
Spelling error
Replace hard-coded references with LaTeX references

@@
 
-Section 2 introduces Elm, signals, state, and the prohibition on signals of
-signals. It is targeted to readers familiar with functional programming but not
-FRP, and may be skipped by those already comfortable with Elm. Section 3
-presents Elm's hover detection for DOM elements, why it initially appears
-inadequate, and how it can be extended to meet our needs. Section 4 explains our
-menu implementation in detail. Section 5 analyzes TodoFRP. Section 6 concludes
-with a notice to the Elm community.
+Section \ref{elm-for-functional-programmers} introduces Elm, signals, state, and
+the prohibition on signals of signals. It is targeted to readers familiar with
+functional programming but not FRP, and may be skipped by those already
+comfortable with Elm.
+Section \ref{detecting-hover-information} presents Elm's hover detection for DOM
+elements, why it initially appears inadequate, and how it can be extended to
+meet our needs.
+Section \ref{implementing-menus} explains our menu implementation in detail.
+Section \ref{related-work-todofrp} analyzes TodoFRP.
+Section \ref{conclusion} concludes with a notice to the Elm community.
 
@@
 (Element, Bool)}. A more mentally convenient type, though impossible to
-obtain,\footnote{We explain why in Section 4.}
+obtain,\footnote{We explain why in Section \ref{implementing-menus}.}
 is \texttt{(Signal Element, Signal Signal Bool)}. There does not exist a general
@@
 \section{Conclusion: To the Elm
-Community}\label{conclusion-to-the-elm-community}
+Community}\label{conclusion}
 
@@
 
-\subsection{Acknowledgements}\label{acknowledgements}
+\subsection{Acknowledgments}\label{acknowledgments}
 

### 2013-12-15 paper/paper.tex
Add citations

@@
 Elm was introduced in March 2012 in Evan Czaplicki's senior thesis,
-\emph{Concurrent FRP for GUIs}. He and his adviser, Stephen Chong,
-published \emph{Asynchronous FRP for GUIs} at PLDI 2013. Both papers are
-comprehensive overviews of Elm, and additionally provide excellent
-literature reviews of previous FRP GUI endeavors, of which there are
-several. Elm compiles down to JavaScript to run in the browser, making
-it remarkably portable.
+\emph{Concurrent FRP for GUIs}. \cite{elm-thesis} He and his adviser, Stephen
+Chong, published \emph{Asynchronous FRP for GUIs} at PLDI
+2013. \cite{CzaplickiC13} Both papers are comprehensive overviews of Elm, and
+additionally provide excellent literature reviews of previous FRP GUI endeavors,
+of which there are several.  Elm compiles down to JavaScript to run in the
+browser, making it remarkably portable.
 
@@
 There are two features of Elm we are deliberately avoiding. First is the
-extensive raster drawing library, \texttt{Graphics.Collage}. Dynamic
-hover detection is not problematic when using this library because it
-can be done purely through manual collision detection. However, if we
-used this library, our GUI would be a single raster animation and not a
-DOM tree. Secondly, the \texttt{Graphics.Input} library contains
-wrappers around HTML checkboxes and dropdowns. We consider creating GUIs
-with these to be uninteresting and wanted to pursue something more
-general. This led to the choice of an interface typically found in the
+extensive raster drawing library, \texttt{Graphics.Collage}. Dynamic hover
+detection is not problematic when using this library because it can be done
+purely through manual collision detection. However, if we used this library, our
+GUI would be a single raster animation and not a DOM tree. Secondly, the
+\texttt{Graphics.Input} library contains wrappers around HTML checkboxes and
+dropdowns. We rejected creating GUIs with these and wanted to pursue something
+more general. This led to the choice of an interface typically found in the
 operating system rather than the browser. We do refer to the native GUI
-constructs to show how the technique we develop for hovering
-generalizes, as the API for buttons and hover detection is very similar.
+constructs to show how the technique we develop for hovering generalizes, as the
+API for buttons and hover detection is very similar.
 
@@
 that (through the use of \texttt{foldp}) depends on the history of an
-input signal, say \texttt{Window.width}.
+input signal, say \texttt{Window.width}. \cite{CzaplickiC13}
 \end{quote}
@@
 when they were created. We avoid these issues by ruling out signals of
-signals.
+signals. \cite{CzaplickiC13}
 \end{quote}
@@
 dimensions are taken to be the maximum of the three inputs' dimensions.
-Even if the Elements have different sizes, the resulting element and
+Even if the Elements have different sizes, the resulting Element and
 therefore the hover surface remains fixed in size. Although the same
@@
 %Note to Max: I think it does work...
+%We should try it out then.
 
@@
 
-\subsection{Acknowledgments}\label{acknowledgments}
-
+\section*{Acknowledgments}
 We would like to thank Evan Czaplicki and Stephen Chong for creating
@@
 
-\subsection{References}\label{references}
 %Make sure to put a link to our GitHub
-
+\bibliographystyle{alpha}
+\bibliography{refs.bib}
 \end{document}

### 2013-12-15 paper/compile.sh
compile script

@@ -1,0 +1,4 @@
+pdflatex paper
+bibtex paper
+pdflatex paper
+bibtex paper

### 2013-12-15 paper/paper.tex
Cut rambling para, add transition

@@
 
-\texttt{Graphics.Input} does not currently support radio buttons, but one could
-imagine them being added in a way similar to the others. However, it is worth
-observing that the pattern itself leaves something to be desired, namely, that
-the functions are impure. One can apply \texttt{hoverable} to an Element twice,
-place the resulting Elements in in different places on the screen (or place one
-not at all), and the signals of Booleans will be different. This is true,
-analogously, of all the functions of similar type in the library. We know of no
-alternative that has been proposed, and do not have one ourselves. However, we
-point out this deficiency in order to illustrate that the library still has room
-for incremental and perhaps even radical growth.
-
 \section{Implementing Menus}\label{implementing-menus}
 
-A menu can be thought of as an instance of a tree.
+Having established the need for \texttt{hoverablesJoin} and implementing it, we
+now turn to stated task of creating menus. A menu can be thought of as an
+instance of a tree.
 

### 2013-12-15 paper/paper.tex
Redid section 4 (explaining our implementation of menus). It flows a lot better now.

@@
 
-Using \texttt{hoverablesJoin}, we created a structure of type
-\texttt{Tree (Signal Element, Signal Bool)} to represent our menus.
-Using the combinator
-\texttt{combine : {[}Signal a{]} -\textgreater{} Signal {[}a{]}}, we
-were able to turn this structure into a
-\texttt{Signal (Tree (Element, Bool))}. Once we had that structure, we
-could render the menu using a pure function from
-\texttt{Tree (Element, Bool) -\textgreater{} Element} and then lift it.
-Similarly to limiting the use of the IO monad in Haskell code, it is
-advantageous to limit the use of signals in Elm code.
+Using \texttt{plainText : String -> Element}, we can take a menu specification
+of type \texttt{Tree (Signal String)} and turn into a \texttt{Tree (Signal
+Element}. From there, we can use \texttt{hoverablesJoin} to get the hover
+information of each element and create a structure of type \texttt{Tree (Signal
+Element, Signal Bool)} to represent our menus.  By recursing on the combinator
+\texttt{combine : {[}Signal a{]} -\textgreater{} Signal {[}a{]}}, we are able to
+turn this structure into a \texttt{Signal (Tree (Element, Bool))}. By changing
+our internal structure into a signal of trees instead of a tree of signals, we
+can do the rest of our work with pure functions that operate on the tree, rather
+than on individual signals.  For example, our rendering function
+\texttt{renderMenu : Tree (Element, Bool) -\textgreater{} Element} can simply be
+lifted onto the menu structure.  Similarly to limiting the use of the IO monad
+in Haskell code, it is advantageous to limit the use of signals in Elm code.
 
-Our first trick was to replace submenus not shown on the screen with
-spacers. A spacer is simply a blank rectangle with a specified with and
+The first problem in implementing menus is twofold: how to position submenus on the
+screen and how to hide submenus that should not be rendered. To get around these
+problems, we replace submenus not shown on the screen with
+spacers. A spacer is simply a blank rectangle with a specified width and
 height
@@
 
-Our second trick fixed a race condition. A submenu was rendered to the
-screen if one of three conditions was met: the mouse was hovering upon
-its parent, the mouse was hovering upon it, or the mouse was hovering
-upon any of its descendants. As the mouse moved from the parent to the
-child, the child was often replaced with a spacer before it could detect
-that the mouse was hovering over it. To fix this problem, we created the
-following function:
+\begin{verbatim}
+elemOrSpacer : (Element, Bool) -> Element
+elemOrSpacer (elem, hover) =
+    if hover
+    then elem
+    else spacer (widthOf elem) 1
+\end{verbatim}
+
+By replacing hidden elements with spacers, we are able to use the \texttt{flow}
+primitive to position our elements instead of manually doing ``absolute''
+positioning.  A submenu is rendered to the screen if one of three conditions
+is met: the mouse is hovering upon its parent, the mouse is hovering upon it,
+or the mouse is hovering upon any of its descendants.
+
+While this allows recursive submenus to be correctly rendered to the screen, it
+creates a race condition. As the mouse moves from the parent to the its submenu,
+the child is often replaced with a spacer before it can detect that the mouse is
+hovering over it. The results in menus disappearing as the mouse changes which
+element it is hovering upon. To fix this problem, we created the following
+function:
 
@@
 
-Using \texttt{extend} makes a \texttt{Signal Bool} wait a millisecond to
+\texttt{extend} makes a \texttt{Signal Bool} wait a millisecond to
 transition from True to False. By applying this function to every
-hovering Boolean in the Menu structure, we gave the mouse time to move
-from a menu Element to its submenu before that submenu disappeared.
+hovering Boolean in the Menu structure, we give the mouse time to move
+from a menu Element to its submenu before that submenu disappears.
+
+A menu that does not detect mouse clicks is a fairly useless menu. There is no
+primitive in Elm that can simply attach click detection to an element. The
+closest primitive is \texttt{customButtons}. However, the implementation of
+\texttt{customButtons} creates an extra dummy spacer which is rendered on top of
+the element and therefore prevents us from changing the color of the menu
+element when the mouse is hovering over it. To get around this issue, we used a
+combination of \texttt{Mouse.clicks : Signal ()} (which fires an event whenever
+the mouse clicks anywhere on the screen) and the hoverable information to
+determine which element (if any) the mouse was hovering over.
+
+\begin{verbatim}
+clicksFromHover : Signal Bool -> Signal ()
+clicksFromHover hover =
+    lift (\_ -> ()) <| keepIf id False <| sampleOn Mouse.clicks hover
+\end{verbatim}
+
+For each menu element in the tree, we found its ``path'' (with type
+\texttt{[Int]}) which denotes the list indexes needed to traverse to that
+element. Combining this with the click information gave each element a signal of
+its path that fired whenever the user clicked on that element.  We then combined
+all of those signals into one signal using the \texttt{merge} combinator
+(\texttt{merges : [Signal a] -> Signal a}). The user of the menu could then use
+this \texttt{Signal [Int]} to respond to mouse clicks. This is significantly
+more flexible than returning a \texttt{Signal String} because a path is
+guaranteed to be unique, whereas a menu string might have a duplicate.
 
@@
 
-Note that the input to our menu is a \texttt{Tree (Signal String)}, but
+Also note that the input to our menu is a \texttt{Tree (Signal String)}, but
 not a \texttt{Signal (Tree String)}. We are able to map from

### 2013-12-15 paper/paper.tex
Revise section 4. Not done yet.

@@
 Having established the need for \texttt{hoverablesJoin} and implementing it, we
-now turn to stated task of creating menus. A menu can be thought of as an
+now turn to stated task of creating menus.
+\subsection{Menu Representations}
+
+A menu specification, the abstract model of a menu, can be thought of as an
 instance of a tree.
@@
 
-Using \texttt{plainText : String -> Element}, we can take a menu specification
-of type \texttt{Tree (Signal String)} and turn into a \texttt{Tree (Signal
-Element}. From there, we can use \texttt{hoverablesJoin} to get the hover
-information of each element and create a structure of type \texttt{Tree (Signal
-Element, Signal Bool)} to represent our menus.  By recursing on the combinator
-\texttt{combine : {[}Signal a{]} -\textgreater{} Signal {[}a{]}}, we are able to
-turn this structure into a \texttt{Signal (Tree (Element, Bool))}. By changing
-our internal structure into a signal of trees instead of a tree of signals, we
-can do the rest of our work with pure functions that operate on the tree, rather
-than on individual signals.  For example, our rendering function
-\texttt{renderMenu : Tree (Element, Bool) -\textgreater{} Element} can simply be
-lifted onto the menu structure.  Similarly to limiting the use of the IO monad
-in Haskell code, it is advantageous to limit the use of signals in Elm code.
+The obvious type for a menu structure that may change is \texttt{Signal (Tree
+String)}. It is easy enough to convert a value of this type to one of type
+\texttt{Signal (Tree Element)}, but we cannot add hover detection to this tree.
+We are able to map from \texttt{Signal Element -\textgreater{} (Signal Element,
+Signal Bool)}, but not (in any useful way) from \texttt{Element -\textgreater{}
+(Element, Bool)}. That is, because the tree is already a signal, we cannot
+create signals inside it (without creating signals of signals). Hover
+information requires a (distinct) signal at each item in the tree.
 
-The first problem in implementing menus is twofold: how to position submenus on the
-screen and how to hide submenus that should not be rendered. To get around these
-problems, we replace submenus not shown on the screen with
-spacers. A spacer is simply a blank rectangle with a specified width and
-height
-(\texttt{spacer : Int -\textgreater{} Int -\textgreater{} Element}). For
-every menu item, we create both the menu Element and a spacer. The
-spacer has the same width as the Element, but a height of 1. We then
-switch between menu and the spacer based on hover information. This
-helped in two ways. First of all, it allowed us to choose between two
-options without creating a signal of signals. Secondly, the spacer
-helped to align the submenu with its parent element.
+Instead of a signal of trees, we therefore use a tree of signals. Each
+individual menu element can contain dynamic information, but the menu structure
+must be static. We can, in practice, get around this restriction by creating a
+tree that is larger than necessary and filling the unused nodes with empty
+string, which our implementation handles appropriately.
+
+Note that we can (and will) convert a tree of signals into a signal of trees, it
+is impossible to go the other way. When converting many signals into one, the
+information on which signal changed is lost. The combinator \texttt{combine :
+    {[}Signal a{]} -\textgreater{} Signal {[}a{]}} does this, and will be used
+    below. It has no inverse because such a function would need to create
+    information: which signal in the list updated? What if the list
+    changes length? What if there is a repeat event, where the list does not
+    change, which signals should propagate repeats? Similar arguments show why
+    the mentally convenient type of \texttt{lift hoverable}, which turned a
+    signal of pairs into a pair of signals, cannot actually be obtained.
+
+\subsection{Menu Rendering}
+
+Therefore, a menu is specified by a value of type \texttt{Tree (Signal String)}.
+We can lift and map the combinator \texttt{plainText : String -> Element} to
+obtain from a menu specification a value of type \texttt{Tree (Signal Element)}.
+From there, we can use \texttt{hoverablesJoin} to get the hover information of
+each Element and create a value of type \texttt{Tree (Signal Element, Signal
+Bool)} to represent our menus. \emph{Now} we convert (by structural recursion)
+our tree of signals to a signal of trees, a value of type \texttt{Signal (Tree
+(Element, Bool))}.  We do the rest of our work with pure functions lifted to
+operate on the tree, for example, our rendering function \texttt{renderMenu :
+    Tree (Element, Bool) -\textgreater{} Element}. Similarly to limiting the use
+    of the IO monad in Haskell code, it is advantageous to limit the use of
+    signals in Elm code.
+
+The first problem in implementing menus is twofold: how to position submenus on
+the screen and how to hide submenus that should not be rendered. To get around
+these problems, we replace submenus not shown on the screen with spacers. A
+spacer is simply a blank rectangle with a specified width and height
+(\texttt{spacer : Int -\textgreater{} Int -\textgreater{} Element}). For every
+menu item, we create both the menu Element and a spacer. The spacer has the same
+width as the Element, but a height of 1. We then switch between menu and the
+spacer based on hover information. This helped in two ways. First of all, it
+allowed us to choose between two options without creating a signal of signals.
+Secondly, the spacer helped to align the submenu with its parent element.
 
@@
 elemOrSpacer (elem, hover) =
-    if hover
-    then elem
+    if hover then elem
     else spacer (widthOf elem) 1
@@
 
-Also note that the input to our menu is a \texttt{Tree (Signal String)}, but
-not a \texttt{Signal (Tree String)}. We are able to map from
-\texttt{Signal Element -\textgreater{} (Signal Element, Signal Bool)},
-but not from \texttt{Element -\textgreater{} (Element, Bool)}.
-Therefore, if we tried to turn a \texttt{Signal (Tree String)} into a
-\texttt{Signal (Tree (Element, Bool))}, we would need to lift an
-impossible hoverable function of type
-\texttt{Element -\textgreater{} (Element, Bool)} on each element.
-Therefore, each individual menu element can contain dynamic information,
-but the menu structure must be static. We can, in practice, get around
-this restriction by creating a tree that is larger than necessary and
-filling the unused nodes with empty string, which our implementation
-handles appropriately.
-
-A possible solution for that issue would be to write a function of type
-\texttt{Signal (Tree a) -\textgreater{} Tree (Signal a)}. Although the
-\texttt{combine} combinator can turn a list of signals into a signal of
-lists, there isn't an inverse to turn a signal of lists into a list of
-signals. It is impossible to write the function
-\texttt{split : Signal {[}a{]} -\textgreater{} {[}Signal a{]}} because
-the size of the list in the input is dynamic, whereas the size of the
-output list is static. Since the \texttt{Tree} data structure contains
-lists, it is therefore impossible to write a function of type
-\texttt{Signal (Tree a) -\textgreater{} Tree (Signal a)}.
-
 \section{Related Work: TodoFRP}\label{related-work-todofrp}
@@
 We present these techniques and analysis in the hopes they are of
-service to the Elm and FRP communities. We hope this paper becomes
+service to the Elm and FRP communities. There are a number of obvious paths that
+dead-end, and non-obvious paths that are fruitful. We hope this paper becomes
 useful in Elm's goal of making GUIs simpler to implement and more robust

### 2013-12-15 paper/paper.tex
More revisions to 4

@@
 
-The first problem in implementing menus is twofold: how to position submenus on
-the screen and how to hide submenus that should not be rendered. To get around
-these problems, we replace submenus not shown on the screen with spacers. A
-spacer is simply a blank rectangle with a specified width and height
-(\texttt{spacer : Int -\textgreater{} Int -\textgreater{} Element}). For every
-menu item, we create both the menu Element and a spacer. The spacer has the same
-width as the Element, but a height of 1. We then switch between menu and the
-spacer based on hover information. This helped in two ways. First of all, it
-allowed us to choose between two options without creating a signal of signals.
-Secondly, the spacer helped to align the submenu with its parent element.
+Because \texttt{hoverablesJoin} was applied throughout the entire tree, it
+contains every Element in every menu, and Booleans that indicate which ones to
+display. Rather than omitting Elements associated with false, we replace these
+hidden submenus with a spacer. A spacer is simply a blank rectangle with a
+specified width and height (\texttt{spacer : Int -\textgreater{} Int
+-\textgreater{} Element}). For our purposes, the spacer has the same width as
+the Element, but a height of 1.\footnote{We were concerned a height of zero
+would cause browser rendering issues.} We then switch between menu and the
+spacer based on hover information. The spacer helped to align the submenu with
+its parent element in a way a \texttt{Maybe Element} would not have done. By
+replacing hidden elements with spacers, we are able to use the \texttt{flow}
+primitive to position our elements instead of manually doing ``absolute''
+positioning.
 
@@
 
-By replacing hidden elements with spacers, we are able to use the \texttt{flow}
-primitive to position our elements instead of manually doing ``absolute''
-positioning.  A submenu is rendered to the screen if one of three conditions
+A submenu is rendered to the screen if one of three conditions
 is met: the mouse is hovering upon its parent, the mouse is hovering upon it,
@@
 
-\texttt{extend} makes a \texttt{Signal Bool} wait a millisecond to
-transition from True to False. By applying this function to every
-hovering Boolean in the Menu structure, we give the mouse time to move
-from a menu Element to its submenu before that submenu disappears.
+Applying \texttt{extend} makes a \texttt{Signal Bool} wait a millisecond to
+transition from true to false. Unlike a delay, true-to-false transitions still
+occur immediately. By applying this function to every hovering Boolean in the
+Menu structure, we give the mouse time to move from a menu Element to its
+submenu before that submenu disappears. A millisecond was all that was needed.
 
 A menu that does not detect mouse clicks is a fairly useless menu. There is no
-primitive in Elm that can simply attach click detection to an element. The
-closest primitive is \texttt{customButtons}. However, the implementation of
-\texttt{customButtons} creates an extra dummy spacer which is rendered on top of
-the element and therefore prevents us from changing the color of the menu
-element when the mouse is hovering over it. To get around this issue, we used a
-combination of \texttt{Mouse.clicks : Signal ()} (which fires an event whenever
-the mouse clicks anywhere on the screen) and the hoverable information to
-determine which element (if any) the mouse was hovering over.
+primitive in Elm that can simply attach click detection to an element. To get
+around this issue, we used a combination of \texttt{Mouse.clicks : Signal ()}
+(which fires an event whenever the mouse clicks anywhere on the screen) and the
+hoverable information to determine which element (if any) the mouse was hovering
+over. (This implementation uses Elm's infix function application operator
+\texttt{|>}, but it can be read as a pipeline.)
 
@@
 clicksFromHover hover =
-    lift (\_ -> ()) <| keepIf id False <| sampleOn Mouse.clicks hover
+    sampleOn Mouse.clicks hover
+    |> keepIf id False
+    |> lift (\_ -> ())
 \end{verbatim}
@@
 
-Interestingly enough, we were able to implement menus without explicitly
-storing the state of the menu. There are two ways to remember state in
-Elm code: \texttt{foldp}, or Elm's implementation of Arrowized FRP known
-as \texttt{Automaton}. However, in our implementation we use the current
-browser DOM state to remember our state. Instead of storing the menu
-currently being displayed on the screen, we can derive it based on which
-Element (if any) the mouse is currently hovering over. The elimination
-of state from our code allows it to be much cleaner.
+Interestingly, we were able to implement menus without explicitly storing the
+state of the menu, i.e. without using \texttt{foldp} (or Elm's implementation of
+Arrowized FRP known as \texttt{Automaton}). Rather we use the current browser
+DOM state to remember state. Instead of storing the menu currently being
+displayed on the screen, we can derive it based on which Element (if any) the
+mouse is currently hovering over. The elimination of state from our code allows
+it to be much cleaner.
 

### 2013-12-15 paper/paper.tex
Fixed a small error
Prefer <| over |> since <| is the same as the Haskell $ operator

@@
 Applying \texttt{extend} makes a \texttt{Signal Bool} wait a millisecond to
-transition from true to false. Unlike a delay, true-to-false transitions still
+transition from true to false. Unlike a delay, false-to-true transitions still
 occur immediately. By applying this function to every hovering Boolean in the
@@
 over. (This implementation uses Elm's infix function application operator
-\texttt{|>}, but it can be read as a pipeline.)
+\texttt{<|}, which is the same as the Haskell operator \texttt{\$}.
 
@@
 clicksFromHover hover =
-    sampleOn Mouse.clicks hover
-    |> keepIf id False
-    |> lift (\_ -> ())
+    lift (\_ -> ())
+    <| keepIf id False
+    <| sampleOn Mouse.clicks hover
 \end{verbatim}

### 2013-12-15 paper/paper.tex
Minor revisions

@@
 
-Elm was introduced in March 2012 in Evan Czaplicki's senior thesis,
-\emph{Concurrent FRP for GUIs}. \cite{elm-thesis} He and his adviser, Stephen
-Chong, published \emph{Asynchronous FRP for GUIs} at PLDI
-2013. \cite{CzaplickiC13} Both papers are comprehensive overviews of Elm, and
-additionally provide excellent literature reviews of previous FRP GUI endeavors,
-of which there are several.  Elm compiles down to JavaScript to run in the
-browser, making it remarkably portable.
+Elm was introduced in March 2012 in Evan Czaplicki's senior thesis.
+\cite{elm-thesis} He and his adviser, Stephen Chong, published a formal
+description of Elm's semantics at PLDI 2013. \cite{CzaplickiC13} Both papers are
+comprehensive overviews of Elm, and additionally provide excellent literature
+reviews of previous FRP GUI endeavors, of which there are several. Elm compiles
+down to JavaScript to run in the browser, making it remarkably portable.
 
-We chose menus as an example GUI to implement in Elm. The particular
-design places the top-level menu at the top of the screen, similar to
-Mac OS X and many Linux distributions, with selections coming down from
-the top. Menus extend when the top-level item is hovered upon, and
-remain extended while the mouse hovers over any item in the menu.
-Therefore it is necessary to know hover information about each menu
-item. This time-varying information is also used to detect selections
-upon click and highlight the moused-over item. It is simple to do this
-when the hover-detecting area is constant. Our paper describes the much
-more difficult task of managing time-varying hover information about
-time-varying areas.
+Elm was designed to make implementing GUIs easier, so we decided to try to
+implement menus as an example GUI. The particular design places the top-level
+menu at the top of the screen, similar to Mac OS X and many Linux distributions,
+with selections coming down from the top. Menus extend when the top-level item
+is hovered upon, and remain extended while the mouse hovers over any item in the
+menu. Therefore it is necessary to know hover information about each menu item.
+This time-varying information is also used to detect selections upon click and
+highlight the moused-over item. It is simple to do this when the hover-detecting
+area is constant. This paper describes the much more difficult task of managing
+time-varying hover information about time-varying areas.
 
@@
 detection is not problematic when using this library because it can be done
-purely through manual collision detection. However, if we used this library, our
+purely using geometric collision detection. However, if we used this library, our
 GUI would be a single raster animation and not a DOM tree. Secondly, the
@@
 operating system rather than the browser. We do refer to the native GUI
-constructs to show how the technique we develop for hovering generalizes, as the
-API for buttons and hover detection is very similar.
+constructs to show how the technique we develop for hovering generalizes to
+them, as their API is similar to the one for hover detection.
 
@@
   An implementation of desktop-style menu in Elm, which incorporates
-  several noteworthy ``tricks'' as well as general style practices.
+  several noteworthy ``tricks'' as well as general style best practices.
 \item
@@
 functional programming but not FRP, and may be skipped by those already
-comfortable with Elm.
-Section \ref{detecting-hover-information} presents Elm's hover detection for DOM
-elements, why it initially appears inadequate, and how it can be extended to
-meet our needs.
-Section \ref{implementing-menus} explains our menu implementation in detail.
-Section \ref{related-work-todofrp} analyzes TodoFRP.
-Section \ref{conclusion} concludes with a notice to the Elm community.
+comfortable with Elm.  Section \ref{detecting-hover-information} presents Elm's
+hover detection API for DOM elements, why it initially appears inadequate, and
+how it can be extended to meet our needs. Section \ref{implementing-menus}
+explains our menu implementation in detail. Section \ref{related-work-todofrp}
+analyzes TodoFRP. Section \ref{conclusion} concludes with a notice to the Elm
+community.
 
@@
 We can print the current area to the screen with \texttt{main = lift asText
-area}. The primitive \texttt{asText : a -\textgreater{} Element} renders almost
+area}. The primitive \texttt{asText : a -> Element} renders almost
 anything into an Element, which represents a DOM element.\footnote{Those
@@
 signals who depend on that signal and the outputs of those signals are
-reevaluated.
+reevaluated, recursively.
 
@@
 The ability to create a signal dependent on past state has a potentially
-disastrous implication for space performance. Czaplicki and Chong
-explain the preventive measure and the problem:
+disastrous implication for space performance. Czaplicki and Chong explain:
 
 \begin{quote}
-The type system rules out programs that use signals of signals, for the
-following reason. Intuitively, if we have signals of signals, then after
-a program has executed for, say 10 minutes, we might create a signal
-that (through the use of \texttt{foldp}) depends on the history of an
-input signal, say \texttt{Window.width}. \cite{CzaplickiC13}
+Intuitively, if we have signals of signals, then after a program has executed
+for, say 10 minutes, we might create a signal that (through the use of
+\texttt{foldp}) depends on the history of an input signal, say
+\texttt{Window.width}. \cite{CzaplickiC13}
 \end{quote}
@@
 
-We could, were the type system not principled on disallowing it, create
+We could, in theory, create
 
@@
 
-That is, in order to have both state and signals of signals, the program
-must either remember indefinitely every event to ever occur, so that a
-newly created signal can use them, or tolerate signals that vary based
-only on when they were created, losing referential transparency. Elm's
-creators declined either option and disallowed signals of signals.
+Indeed, Elm's type system is principled on disallowing signals of signals.
+Programs may not create in the future signals that depend on the past. Were they
+allowed to, the system must either remember indefinitely every event to ever
+occur on any signal, so that a newly created signal can use them, or tolerate
+signals that vary based only on when they were created, losing referential
+transparency.
 
@@
 
-\texttt{hoverable : Element -\textgreater{} (Element, Signal Bool)}
+\texttt{hoverable : Element -> (Element, Signal Bool)}
 
-The returned Element is visually identical to argument, but now detects
+The returned Element is visually identical to the argument, but now detects
 hover information. The signal of Booleans reflects the hover status of
@@
 
-\texttt{lift hoverable : Signal Element -\textgreater{} Signal (Element, Signal Bool)}
+\texttt{lift hoverable : Signal Element -> Signal (Element, Signal Bool)}
 
@@
 (Element, Bool)}. A more mentally convenient type, though impossible to
-obtain,\footnote{We explain why in Section \ref{implementing-menus}.}
+obtain,\footnote{We explain why in Section \ref{menu-representations}.}
 is \texttt{(Signal Element, Signal Signal Bool)}. There does not exist a general
@@
 has clicked on an Element so that the system may respond, we must know which one
-is being hovered over.\footnote{Once this is known, \texttt{sampleOn
-Mouse.clicks} can be applied to obtain a signal of the hovered-over element
-updated on each click.}
+is being hovered over.
 
@@
 
-\texttt{hoverableJoin: Signal Element -\textgreater{} (Signal Element, Signal Bool)}
+\texttt{hoverableJoin: Signal Element -> (Signal Element, Signal Bool)}
 
-Comparing its type to the mentally convenient type of \texttt{lift hoverables},
+Comparing its type to the mentally convenient type of \texttt{lift hoverable},
 we see it is identical except that the Boolean value is a signal, not a signal
-of signals. This function is implemented using the more general
-\texttt{hoverables} (note the plural) primitive, of the following type:
+of signals. That is, it has been joined.\footnote{While the name
+    \texttt{hoverableJoin} reinforces this fact for our paper, we do not
+recommend it for a user-facing API.} It is implemented using the more general
+primitive \texttt{hoverables} (note the plural), of the following type:
 
@@
 be defined and so a default value is required). The returned record
-includes the \texttt{events} signals and the \texttt{hoverable}
+includes the \texttt{events} signal and the \texttt{hoverable}
 function, which in general may be applied multiple times so that
-multiple elements report on \texttt{events}. The
-\texttt{(Bool -\textgreater{} a)} is used to identify which Element
+multiple elements report their hover information on \texttt{events}. The
+\texttt{(Bool -> a)} is used to identify which Element
 experienced the event; a common use is \texttt{a = (Int, Bool)} where
@@
 \begin{verbatim}
-hoverablesJoin : Signal Element -> (Signal Element, Signal Bool)
-hoverablesJoin elem =
+hoverableJoin : Signal Element -> (Signal Element, Signal Bool)
+hoverableJoin elem =
     let pool = hoverables False
@@
 
-When a element changes, \texttt{hoverablesJoin} attaches the same
-hovering signal to the new element. If we were to display both the old
-and the new elements on the screen, hovering over either would trigger
-the signal.
+When a element changes, \texttt{hoverableJoin} attaches the new element to the
+same hovering signal. If we were to display both the old and the new elements on
+the screen, hovering over either would trigger the signal.
 
@@
 Boolean state so dynamic behavior has limited use. However, we can
-imagine the ability to disable (``gray-out'') a checkbox in response to
-other choices in a form.
+imagine the ability to dynamically disable (``gray-out'') a checkbox in response
+to other choices in a form.
 
@@
 only because there seems to be no way to supply a
-\texttt{Signal {[}String{]}}, but because drop-downs are commonly used
+\texttt{Signal {[}String{]}}, but also because drop-downs are commonly used
 in groups. Multiple menus reporting on one signal would be very
-convenient. If the menus were dynamic, it would be useful for the API to
-indicate when an event occurred because the user makes a selection vs.~when the
+convenient. If the menu options were dynamic, it would be useful for the API to
+indicate when an event occurred because the user made a selection vs.~when the
 menu updated and the previously selected string was no longer available.
@@
 
-Having established the need for \texttt{hoverablesJoin} and implementing it, we
+Having established the need for \texttt{hoverableJoin} and implementing it, we
 now turn to stated task of creating menus.
-\subsection{Menu Representations}
+\subsection{Menu Representations}\label{menu-representations}
 
@@
 \texttt{Signal (Tree Element)}, but we cannot add hover detection to this tree.
-We are able to map from \texttt{Signal Element -\textgreater{} (Signal Element,
-Signal Bool)}, but not (in any useful way) from \texttt{Element -\textgreater{}
+We are able to map from \texttt{Signal Element -> (Signal Element,
+Signal Bool)}, but not (in any useful way) from \texttt{Element ->
 (Element, Bool)}. That is, because the tree is already a signal, we cannot
@@
 
-Note that we can (and will) convert a tree of signals into a signal of trees, it
-is impossible to go the other way. When converting many signals into one, the
-information on which signal changed is lost. The combinator \texttt{combine :
-    {[}Signal a{]} -\textgreater{} Signal {[}a{]}} does this, and will be used
-    below. It has no inverse because such a function would need to create
-    information: which signal in the list updated? What if the list
-    changes length? What if there is a repeat event, where the list does not
-    change, which signals should propagate repeats? Similar arguments show why
-    the mentally convenient type of \texttt{lift hoverable}, which turned a
-    signal of pairs into a pair of signals, cannot actually be obtained.
+Note that we can (and will) convert a tree of signals into a signal of trees,
+but it is impossible to go the other way. When converting many signals into one,
+the information on which signal changed is lost. The combinator \texttt{combine
+    : {[}Signal a{]} -> Signal {[}a{]}} does this, and will be used below. It
+    has no inverse because such a function would need to create information:
+    which signal in the list updated? What if the list changes length? What if
+    there is a repeat event, where the list does not change, which signals
+    should propagate repeats? Similar arguments show why the mentally convenient
+    type of \texttt{lift hoverable}, which turned a signal of pairs into a pair
+    of signals, cannot actually be obtained.
 
@@
 obtain from a menu specification a value of type \texttt{Tree (Signal Element)}.
-From there, we can use \texttt{hoverablesJoin} to get the hover information of
+From there, we can use \texttt{hoverableJoin} to get the hover information of
 each Element and create a value of type \texttt{Tree (Signal Element, Signal
@@
 operate on the tree, for example, our rendering function \texttt{renderMenu :
-    Tree (Element, Bool) -\textgreater{} Element}. Similarly to limiting the use
+    Tree (Element, Bool) -> Element}. Similarly to limiting the use
     of the IO monad in Haskell code, it is advantageous to limit the use of
@@
 
-Because \texttt{hoverablesJoin} was applied throughout the entire tree, it
+Because \texttt{hoverableJoin} was applied throughout the entire tree, it
 contains every Element in every menu, and Booleans that indicate which ones to
@@
 hidden submenus with a spacer. A spacer is simply a blank rectangle with a
-specified width and height (\texttt{spacer : Int -\textgreater{} Int
--\textgreater{} Element}). For our purposes, the spacer has the same width as
+specified width and height (\texttt{spacer : Int -> Int
+-> Element}). For our purposes, the spacer has the same width as
 the Element, but a height of 1.\footnote{We were concerned a height of zero
 would cause browser rendering issues.} We then switch between menu and the
-spacer based on hover information. The spacer helped to align the submenu with
+spacer based on hover information. The spacer helps to align the submenu with
 its parent element in a way a \texttt{Maybe Element} would not have done. By
@@
 
-A submenu is rendered to the screen if one of three conditions
-is met: the mouse is hovering upon its parent, the mouse is hovering upon it,
-or the mouse is hovering upon any of its descendants.
-
-While this allows recursive submenus to be correctly rendered to the screen, it
-creates a race condition. As the mouse moves from the parent to the its submenu,
-the child is often replaced with a spacer before it can detect that the mouse is
-hovering over it. The results in menus disappearing as the mouse changes which
-element it is hovering upon. To fix this problem, we created the following
-function:
+There was another complication involving the hover information. A submenu is
+rendered to the screen if one of three conditions is met: the mouse is hovering
+upon its parent, the mouse is hovering upon it, or the mouse is hovering upon
+any of its descendants. While this definition allows recursive submenus to be
+correctly rendered to the screen, it creates a race condition. As the mouse
+moves from the parent to the its submenu, the child is often replaced with a
+spacer before it can detect that the mouse is hovering over it. The results in
+menus disappearing as the mouse changes which element it is hovering upon. To
+fix this problem, we created the following function:
 
 \begin{verbatim}
-extend : Signal Bool -> Signal Bool
-extend b = lift2 (||) b (delay millisecond b)
+delayFalse : Signal Bool -> Signal Bool
+delayFalse b = lift2 (||) b (delay millisecond b)
 \end{verbatim}
 
-Applying \texttt{extend} makes a \texttt{Signal Bool} wait a millisecond to
-transition from true to false. Unlike a delay, false-to-true transitions still
-occur immediately. By applying this function to every hovering Boolean in the
-Menu structure, we give the mouse time to move from a menu Element to its
-submenu before that submenu disappears. A millisecond was all that was needed.
+Applying \texttt{delayFalse} makes a \texttt{Signal Bool} wait a millisecond to
+transition from true to false. Unlike a typical delay, false-to-true transitions
+still occur immediately. By applying this function to every hovering Boolean in
+the menu structure, we give the mouse time to move from a menu Element to its
+submenu before that submenu disappears.
 
@@
 over. (This implementation uses Elm's infix function application operator
-\texttt{<|}, which is the same as the Haskell operator \texttt{\$}.
+\texttt{<|}, which is the same as the Haskell operator \texttt{\$}).
 
@@
 its path that fired whenever the user clicked on that element. We then combined
-all of those signals into one signal using the \texttt{merge} combinator
+all of those signals into one signal using the \texttt{merges} combinator
 (\texttt{merges : [Signal a] -> Signal a}). The user of the menu could then use
@@
 We examine \href{https://github.com/evancz/TodoFRP}{TodoFRP}, a simple
-Elm web app created by Czapliki, to both pinpoint the problem and show
-its universality. TodoFRP is the current state-of-the-art in highly
+Elm web app created by Czapliki, to provide another example of the problem
+solved by \texttt{hoverableJoin}. TodoFRP is the current state-of-the-art in highly
 reactive Elm GUIs. It provides examples of different levels of
@@
 TodoFRP presents the user with a text field asking, ``what needs to be
-done?''. Entered TODO entries become DOM elements, which can be deleted
+done?''. Entered TODO items become DOM elements, which can be deleted
 with a ``x'' button, also a DOM element. The button is implemented using
@@
 as intended.
-%Note to Max: I think it does work...
-%We should try it out then.
 
-In the case of TodoFRP, these Elements are different colors of the ``x''
+In the case of TodoFRP, these three Elements are different colors of the ``x''
 and the same for each TODO entry. The polymorphic \texttt{a}s are unique
@@
 The button Elements that do detect hover information are known
-statically. Without \texttt{hoverablesJoin}, it would be impossible to
+statically. With \texttt{hoverableJoin}, it becomes possible to
 dismiss a TODO by clicking on its text label.
@@
 Elm, and the Elm community for growing it. We thank Norman Ramsey for
-his guidance through functional programming, and our paper reviewers,
-\ldots{} .
+his guidance through functional programming, and our paper reviewer, Caroline
+Marcks.
 

### 2013-12-15 paper/paper.tex
Decapitalize Element, Boolean, when not in code

@@
 We can print the current area to the screen with \texttt{main = lift asText
-area}. The primitive \texttt{asText : a -> Element} renders almost
-anything into an Element, which represents a DOM element.\footnote{Those
-    following along in an Elm compiler, such as the one available at
+area}. The primitive \texttt{asText : a -> Element} renders almost anything into
+an \texttt{Element}, which represents a DOM element.\footnote{Those following
+    along in an Elm compiler, such as the one available at
     \texttt{elm-lang.org/try}, should add \texttt{import Window} to the top of
@@
 Elm's \texttt{Graphics.Input} library provides a function to obtain the hover
-information from an Element.
+information from an element.
 
@@
 
-The returned Element is visually identical to the argument, but now detects
-hover information. The signal of Booleans reflects the hover status of
-the this Element, not the original. This function works well when the
-Element is pure (not a signal). For example, if the top-level menu items are
-known statically and never change, they can be rendered as pure Elements with
+The returned element is visually identical to the argument, but now detects
+hover information. The signal of booleans reflects the hover status of
+the this element, not the original. This function works well when the
+element is pure (not a signal). For example, if the top-level menu items are
+known statically and never change, they can be rendered as pure elements with
 this function. We can also create a function, perhaps of type \texttt{Bool ->
@@
 the user switches among applications. Finally, in order to determine when a user
-has clicked on an Element so that the system may respond, we must know which one
+has clicked on an element so that the system may respond, we must know which one
 is being hovered over.
@@
 Comparing its type to the mentally convenient type of \texttt{lift hoverable},
-we see it is identical except that the Boolean value is a signal, not a signal
+we see it is identical except that the boolean value is a signal, not a signal
 of signals. That is, it has been joined.\footnote{While the name
@@
 multiple elements report their hover information on \texttt{events}. The
-\texttt{(Bool -> a)} is used to identify which Element
+\texttt{(Bool -> a)} is used to identify which element
 experienced the event; a common use is \texttt{a = (Int, Bool)} where
-the integer identifies the Element and the Bool is the event.
+the integer identifies the element and the boolean is the event.
 Additionally, \texttt{hoverables} is used to implement
@@
 It ignores the polymorphism (\texttt{a = Bool}) and instead create a
-Boolean signal that is originally false and use the identity function to
+boolean signal that is originally false and use the identity function to
 not alter the hoverable information. With a simple change, we can create
@@
 More dangerously, it becomes easy to create an infinite loop. Suppose an
-Element shrinks on hover. Suppose the cursor hovers on the Element,
-which is then replaced by a smaller Element, so the cursor is no longer
-hovering on the Element. Then the original Element is put back, but is
+element shrinks on hover. Suppose the cursor hovers on the element,
+which is then replaced by a smaller element, so the cursor is no longer
+hovering on the element. Then the original element is put back, but is
 now being hovered on! This condition manifests itself as flickering
-between the two Elements. It is unavoidable in any language or system
+between the two elements. It is unavoidable in any language or system
 that allows hover targets to change size in response to hover events.
@@
 program, and checkboxes. As currently implemented, checkboxes have
-Boolean state so dynamic behavior has limited use. However, we can
+boolean state so dynamic behavior has limited use. However, we can
 imagine the ability to dynamically disable (``gray-out'') a checkbox in response
@@
 From there, we can use \texttt{hoverableJoin} to get the hover information of
-each Element and create a value of type \texttt{Tree (Signal Element, Signal
+each element and create a value of type \texttt{Tree (Signal Element, Signal
 Bool)} to represent our menus. \emph{Now} we convert (by structural recursion)
@@
 Because \texttt{hoverableJoin} was applied throughout the entire tree, it
-contains every Element in every menu, and Booleans that indicate which ones to
-display. Rather than omitting Elements associated with false, we replace these
+contains every element in every menu, and booleans that indicate which ones to
+display. Rather than omitting elements associated with false, we replace these
 hidden submenus with a spacer. A spacer is simply a blank rectangle with a
@@
 -> Element}). For our purposes, the spacer has the same width as
-the Element, but a height of 1.\footnote{We were concerned a height of zero
+the element, but a height of 1.\footnote{We were concerned a height of zero
 would cause browser rendering issues.} We then switch between menu and the
@@
 transition from true to false. Unlike a typical delay, false-to-true transitions
-still occur immediately. By applying this function to every hovering Boolean in
-the menu structure, we give the mouse time to move from a menu Element to its
+still occur immediately. By applying this function to every hovering boolean in
+the menu structure, we give the mouse time to move from a menu element to its
 submenu before that submenu disappears.
@@
 DOM state to remember state. Instead of storing the menu currently being
-displayed on the screen, we can derive it based on which Element (if any) the
+displayed on the screen, we can derive it based on which element (if any) the
 mouse is currently hovering over. The elimination of state from our code allows
@@
 \texttt{customButton} provides the identifier event when the button is
-clicked, and three (pure) Elements to display: one normally, one on
+clicked, and three (pure) elements to display: one normally, one on
 hover, and one on click. The result is an \texttt{Element}, not a
 \texttt{Signal Element}, that nevertheless changes among those three in
-response to the mouse. This is possible because the result Element's
+response to the mouse. This is possible because the result element's
 dimensions are taken to be the maximum of the three inputs' dimensions.
-Even if the Elements have different sizes, the resulting Element and
+Even if the elements have different sizes, the resulting element and
 therefore the hover surface remains fixed in size. Although the same
@@
 
-In the case of TodoFRP, these three Elements are different colors of the ``x''
+In the case of TodoFRP, these three elements are different colors of the ``x''
 and the same for each TODO entry. The polymorphic \texttt{a}s are unique
@@
 
-The TODO label Elements are dynamic and do not detect hover information.
-The button Elements that do detect hover information are known
+The TODO label elements are dynamic and do not detect hover information.
+The button elements that do detect hover information are known
 statically. With \texttt{hoverableJoin}, it becomes possible to

### 2013-12-16 paper/paper.tex
add link to github
Cite todoFRP, forum properly

@@
 \href{http://docs.elm-lang.org/library/Graphics/Input.elm\#hoverables}{documentation}
-for \texttt{hoverables} states that it allows users to ``create and
-destroy elements dynamically and still detect hover information,'' but
-gives no further indicators on how to do so. Though the Elm website is
-full of examples, there are none for either \texttt{hoverable} or
-\texttt{hoverables}. Moreover, in the
-\href{https://groups.google.com/d/msg/elm-discuss/QgowLy5jdhA/CZQfjkbjMsEJ}{mailing
-list post} that introduced these functions, Czaplicki said that
-``\texttt{hoverables} is very low level, but the idea is that you can
-build any kind of nicer abstraction on top of it.'' We have done just
+for \texttt{hoverables} states that it allows users to ``create and destroy
+elements dynamically and still detect hover information,'' but gives no further
+indicators on how to do so. Though the Elm website is full of examples, there
+are none for either \texttt{hoverable} or \texttt{hoverables}. Moreover, in the
+mailing list post that introduced these functions, Czaplicki said that
+``\texttt{hoverables} is very low level, but the idea is that you can build any
+kind of nicer abstraction on top of it.'' \cite{hover-forum} We have done just
 that.
@@
 
-We examine \href{https://github.com/evancz/TodoFRP}{TodoFRP}, a simple
-Elm web app created by Czapliki, to provide another example of the problem
+We examine TodoFRP, \cite{todoFRP} a simple
+Elm web app created by Czaplicki, to provide another example of the problem
 solved by \texttt{hoverableJoin}. TodoFRP is the current state-of-the-art in highly
@@
 
+Our code is available at \url{https://github.com/ealter/Elm-GUIs}.
+
 \section*{Acknowledgments}
@@
 
-%Make sure to put a link to our GitHub
 \bibliographystyle{alpha}

### 2014-01-01 paper/paper.tex
Solidify our explanation about why we can't do hoverable information on a signal of trees

@@
 \texttt{Signal (Tree Element)}, but we cannot add hover detection to this tree.
-We are able to map from \texttt{Signal Element -> (Signal Element,
-Signal Bool)}, but not (in any useful way) from \texttt{Element ->
-(Element, Bool)}. That is, because the tree is already a signal, we cannot
-create signals inside it (without creating signals of signals). Hover
-information requires a (distinct) signal at each item in the tree.
+We are able to map from \texttt{Signal Element -> (Signal Element, Signal
+Bool)}, but not (in any useful way) from \texttt{Element -> (Element, Bool)}. We
+obviously cannot create a \texttt{Signal (Tree (Signal Element, Signal Bool))}
+because that would constitute a signal of signals.  Therefore, we would need to
+create a function of type \texttt{Tree String -> Tree (Element, Bool)} and then
+\texttt{lift} it onto the \texttt{Signal (Tree String)}. However, since we
+cannot write the hovering function of type \texttt{Element -> (Element, Bool)},
+this is impossible.  Therefore, we cannot detect hoverable information on the
+\texttt{Signal (Tree String)}. Hover information requires a (distinct) signal at
+each item in the tree.
 

### 2014-01-01 paper/paper.tex
Explain hoverables better

--- 2013-12-16 paper/paper.tex
@@
 
-The polymorphic \texttt{a} type can serve as an identifier. The first
-value supplies the default value of \texttt{events} (signals must always
-be defined and so a default value is required). The returned record
-includes the \texttt{events} signal and the \texttt{hoverable}
-function, which in general may be applied multiple times so that
-multiple elements report their hover information on \texttt{events}. The
-\texttt{(Bool -> a)} is used to identify which element
-experienced the event; a common use is \texttt{a = (Int, Bool)} where
-the integer identifies the element and the boolean is the event.
-Additionally, \texttt{hoverables} is used to implement
-\texttt{hoverable}:
+A typical use of \texttt{hoverables} is as follows. An initial
+\texttt{a} is supplied that becomes the default value of
+\texttt{events}. If \texttt{hoverable} is never applied, \texttt{events}
+is constant. If it is, the \texttt{Element} argument of
+\texttt{hoverable} in the record acts much like it does in
+\texttt{hoverable} the top-level combinator: provides the static visual
+appearance of the result \texttt{Element}, which also has hover
+detection added. Hover events are boolean, so the function argument is
+called on every such event to produce an \texttt{a} that becomes an
+event on \texttt{events}.
+
+What gives \texttt{hoverables} its plural name is that the record's
+\texttt{hoverable} may be applied multiple times so that multiple
+elements report their hover information on \texttt{events}. The
+polymorphic \texttt{a} type exists to uniquely identify each element,
+though it is also common for the \texttt{a} to also include the boolean
+event. The simplest choice of \texttt{a} is simply \texttt{Bool}, and
+this allows \texttt{hoverables} to implement \texttt{hoverable}:
 
@@
 
-It ignores the polymorphism (\texttt{a = Bool}) and instead create a
-boolean signal that is originally false and use the identity function to
-not alter the hoverable information. With a simple change, we can create
-a function that acts on \texttt{Signal Element} instead of
+The \texttt{a -> Bool} function trivially becomes the identity. With a simple
+change, we can create a function that acts on \texttt{Signal Element} instead of
 \texttt{Element}:

### 2014-01-01 paper/paper.tex
Explain hover target resize flickering better

@@
 
-More dangerously, it becomes easy to create an infinite loop. Suppose an
-element shrinks on hover. Suppose the cursor hovers on the element,
-which is then replaced by a smaller element, so the cursor is no longer
-hovering on the element. Then the original element is put back, but is
-now being hovered on! This condition manifests itself as flickering
-between the two elements. It is unavoidable in any language or system
-that allows hover targets to change size in response to hover events.
+More dangerously, it becomes easy to create an infinite loop. Suppose an element
+shrinks on hover and returns to its original size without hover. Now suppose the
+cursor hovers on the element, which is then replaced by a smaller element, so
+the cursor is no longer hovering on the element. Then the original element is
+put back, but is now being hovered on because the mouse has not moved, so the
+cycle repeats. This condition often manifests itself as flickering between the
+two elements. As this description is implementation-independent, it is
+unavoidable in any language or system that allows hover targets to change size
+in response to hover events on that target.
 

### 2014-01-01 paper/paper.tex
More clarifications

@@
   modifications to the Elm compiler or runtime, and that generalizes to
-  other functions in \texttt{Graphics.Input}.
+  other functions in \texttt{Graphics.Input}. Concretely, this is the
+  implementation of \texttt{hoverableJoin}.
 \item
-  An implementation of desktop-style menu in Elm, which incorporates
+  An implementation of desktop-style menus in Elm, which incorporates
   several noteworthy ``tricks'' as well as general style best practices.
@@
 it is hoped that as the community becomes more familiar with functional
-GUIs, new libraries are added to incorporate some of our tricks, or even
+GUIs, new libraries are added that incorporate some of our tricks, or even
 make them unnecessary. Elm's upcoming third-party library sharing system

### 2014-01-01 paper/paper.tex
Clarified hoverables explanation

--- 2014-01-01 paper/paper.tex
@@
 
-A typical use of \texttt{hoverables} is as follows. An initial
-\texttt{a} is supplied that becomes the default value of
-\texttt{events}. If \texttt{hoverable} is never applied, \texttt{events}
-is constant. If it is, the \texttt{Element} argument of
-\texttt{hoverable} in the record acts much like it does in
-\texttt{hoverable} the top-level combinator: provides the static visual
-appearance of the result \texttt{Element}, which also has hover
-detection added. Hover events are boolean, so the function argument is
-called on every such event to produce an \texttt{a} that becomes an
-event on \texttt{events}.
+The first parameter to
+\texttt{hoverables} denotes the initial value of the \texttt{events} signal
+(signals in Elm must always be defined).
+The \texttt{hoverable} part of the record takes an element and attaches hover
+detection to it. Whenever the mouse hovers (or stops hovering) over the
+element, an event is fired on the \texttt{events} signal. The value of that
+event is determined by the \texttt{(Bool -> a)} parameter of the
+\texttt{hoverable} function. In many cases, this will be \texttt{id} so that the
+\texttt{events} signal is \texttt{True} when the element is being hovered upon
+and \texttt{False} when it is not. However, the \texttt{(Bool -> a)} parameter
+allows you to provide a different value on the \texttt{events} signal for different elements.
 
-What gives \texttt{hoverables} its plural name is that the record's
-\texttt{hoverable} may be applied multiple times so that multiple
-elements report their hover information on \texttt{events}. The
-polymorphic \texttt{a} type exists to uniquely identify each element,
-though it is also common for the \texttt{a} to also include the boolean
-event. The simplest choice of \texttt{a} is simply \texttt{Bool}, and
-this allows \texttt{hoverables} to implement \texttt{hoverable}:
+The more widely used \texttt{hoverable} function is implemented using
+\texttt{hoverables}.
 
@@
 
-The \texttt{a -> Bool} function trivially becomes the identity. With a simple
-change, we can create a function that acts on \texttt{Signal Element} instead of
-\texttt{Element}:
+There is a subtle, yet very important difference between \texttt{hoverables} and
+\texttt{hoverable}. After providing the initial parameter to
+\texttt{hoverables}, you are given a signal and a way to attach elements to that
+signal. By applying the \texttt{hoverable} part of the record to multiple
+elements, we can create a single signal which responds to the hover information
+of those elements, instead of one signal per element. We can then exploit this
+property to detect hoverable information on a signal of elements. As the value
+of the input signal of elements changes, we attach the hover information to the
+new element without having to create a new output signal.
 
@@
 Notice that \texttt{pool.hoverable} is partially applied to \texttt{id}
-purely, and then lifted on to the argument. This is possible,
-ultimately, because \texttt{pool.hoverable} is pure.
+purely, and then lifted on to the argument.
 

### 2014-01-01 paper/paper.tex
More revisions on hoverables

@@
 
-The first parameter to
-\texttt{hoverables} denotes the initial value of the \texttt{events} signal
-(signals in Elm must always be defined).
-The \texttt{hoverable} part of the record takes an element and attaches hover
-detection to it. Whenever the mouse hovers (or stops hovering) over the
-element, an event is fired on the \texttt{events} signal. The value of that
-event is determined by the \texttt{(Bool -> a)} parameter of the
-\texttt{hoverable} function. In many cases, this will be \texttt{id} so that the
-\texttt{events} signal is \texttt{True} when the element is being hovered upon
-and \texttt{False} when it is not. However, the \texttt{(Bool -> a)} parameter
-allows you to provide a different value on the \texttt{events} signal for different elements.
+The first parameter to \texttt{hoverables} denotes the initial value of the
+\texttt{events} signal (signals in Elm must always be defined). The record's
+\texttt{hoverable} function is applied to an element and attaches hover
+detection to it. Whenever the mouse hovers (or stops hovering) over the element,
+an event is fired on the \texttt{events} signal. The value of that event is
+determined by the \texttt{(Bool -> a)} parameter of the \texttt{hoverable}
+function. In many cases, \texttt{a} will be \texttt{Bool} and the function will
+be \texttt{id}. In general, the record's \texttt{hoverable} may be applied
+multiple times with all elements it produces reporting their hover information
+on \texttt{events}. In this case, \texttt{a} can serve as an identifier, and is
+provided by the function argument.
 
@@
 
-There is a subtle, yet very important difference between \texttt{hoverables} and
-\texttt{hoverable}. After providing the initial parameter to
-\texttt{hoverables}, you are given a signal and a way to attach elements to that
-signal. By applying the \texttt{hoverable} part of the record to multiple
-elements, we can create a single signal which responds to the hover information
-of those elements, instead of one signal per element. We can then exploit this
-property to detect hoverable information on a signal of elements. As the value
-of the input signal of elements changes, we attach the hover information to the
-new element without having to create a new output signal.
+Using only \texttt{hoverable}, we are restricted to one element per signal. With
+\texttt{hoverables}, the hover information of an element can be added to an
+existing signal, and can even be added during program execution, that is, when
+the element is a signal. As the value of a signal of elements changes, we attach
+the hover information to the each new element without having to create a new
+output signal. We do this by partially applying \texttt{pool.hoverable} to
+\texttt{id} purely, and then lifting it on to the argument.
 
@@
 \end{verbatim}
-
-Notice that \texttt{pool.hoverable} is partially applied to \texttt{id}
-purely, and then lifted on to the argument.
 

### 2014-01-01 paper/paper.tex
more revisions

@@
 be \texttt{id}. In general, the record's \texttt{hoverable} may be applied
-multiple times with all elements it produces reporting their hover information
+multiple times, with all elements it produces reporting their hover information
 on \texttt{events}. In this case, \texttt{a} can serve as an identifier, and is
@@
 
-Using only \texttt{hoverable}, we are restricted to one element per signal. With
-\texttt{hoverables}, the hover information of an element can be added to an
-existing signal, and can even be added during program execution, that is, when
-the element is a signal. As the value of a signal of elements changes, we attach
-the hover information to the each new element without having to create a new
-output signal. We do this by partially applying \texttt{pool.hoverable} to
-\texttt{id} purely, and then lifting it on to the argument.
+Using only \texttt{hoverable}, we are restricted to one element per signal. By
+contrast, \texttt{hoverables} permits multiple elements on a single signal. The
+hover information of an element can be added to an existing signal, and can even
+be added dynamically, that is, when the element is a signal. As the value of a
+signal of elements changes, we attach the hover information to the each new
+element without having to create a new output signal. We do this by partially
+applying \texttt{pool.hoverable} to \texttt{id} purely, and then lifting it on
+to the argument.
 
